Appendix part 2

### 2025-11-03 Output/authoritarian_only_word_pop_rev.Rmd
Updates to theory section

Mass protest is one of the two primary threats to authoritarian regimes [@svolik_politics_2012], and as such they often go to great lengths to prevent mass protest. These efforts include costly (and risky) investments in the repressive apparatus [@Escriba_Folch2013;@draguPreventiveRepressionTwo2019], co-optation of opposition parties [@reuter2015], and large-scale monitoring and censorship schemes [@King2013]. And they often relentlessly hound would-be protest entrepreneurs to deter such behavior [@xuRepressCooptAuthoritarian2021a;@mehrlSecretPoliceOrganizations2024]. Mass protest then, is clearly undesirable from the authoritarian incumbent's point of view, and we can plausibly assume that once an electoral protest movement begins, the incumbent will exert considerable effort to quash it. Governments that are stronger, relative to the protesting opposition, will defeat protest movements faster than will weaker governments.

An important objection could be raised here. Might not a stronger incumbent, feeling secure in their power, be willing to tolerate opposition protest longer than a weak incumbent? It is true that incumbents do sometimes tolerate protest movements [@yuenNeitherRepressionConcession2017]. However, there are two reasons this is unlikely to be the case systematically for the circumstances studied here. First, authoritarian governments are more likely to tolerate protests aimed at low-salience, local issues, and more likely to clamp down on protests that threaten the legitimacy of the regime. As electoral protests challenge the incumbent's ostensible victory and right to remain in office, they are directly threatening to the regime. Second, authoritarian leaders are working in an information-poor environment, in which mass protests can unexpectedly grow from small sparks to wildfires that quickly sweep away the regime [@Kuran1991]. We should thus expect incumbents to work to defuse them as quickly as possible.

Talk about protest duration here:

Election manipulation is generally characterized as having legitimacy costs which reduce other actors' level of cooperation with the incumbent [@birch2011]. The most visible, and arguably most studied, form of such legitimacy costs is mass post-election protest against the incumbent. There is a broad consensus in the literature that authoritarian incumbents find sustained, sizable protests undesirable, since they entail some risk of incumbent defeat and may require costly repression or concessions. I assume, as a result, that incumbents generally seek to demobilize protest; protest movements that survive longer indicate the incumbent is relatively weaker compared to the opposition.^[The implications of protest toleration by incumbents will be discussed later.] By contrast, when protest movements meet a quick demise, the incumbent can be considered relatively strong.^[In both cases, this assume that incumbent survives the protest movement, by far the modal outcome. When the opposition wins, an outcome considered in some of the empirical models below, the assumption is reversed: a quicker opposition victory indicates a weaker incumbent.] As argued below, electoral manipulation can be demonstrative of capacities that may also be used to defeat protest movements.

There is an alternative view of electoral manipulation, however; Simpser's foundational argument holds that manipulated elections can send costly signals of the incumbent's "capacities, resources, [and] dispositions" that will help it retain power in the future [@simpser2013, p. 80]---that is, of its political strength. I argue that more recent research on the principal-agent dynamics inherent in election manipulation reinforces and expands this idea. This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver]. Incumbents that can solve these principal-agent problems at election-time, are also more likely to be able to survive post-election challenges.

@@

This argument---that successful, large-scale election manipulation draws on the same resources a state may use to suppress protest---turns the deterrent logic on its head. Instead of signalling weakness, large-scale fraud indicates a well of resilience: control over resources, impunity for pro-regime lawbreaking, and loyal agents. As a result, governments that build and deploy an electoral manipulation apparatus are simultaneously inoculating themselves against the expected cost of protest. This yields Hypothesis 2.

This true level of support can be obscured prior to elections due to pervasive preference falsification [@Kuran1991] and the low level of political information common in authoritarian regimes [@guriev2020].  <- Keep this for risk of protests spiraling out of control

Regime-weakness models of electoral manipulation, discussed above, treat election manipulation and genuine votes as substitute goods. For any given vote-share, an increase in manufactured votes reduces the number of genuine votes. In these models, which consider manufactured votes to be mere illusions of strength, the higher the number of manufactured votes for any given vote total, the weaker the incumbent. This means that strategic opposition actors are more likely to initiate protest when they perceive the incumbent to be weaker, and expect to be more likely to succeed in a challenge. In other words, manipulated votes are empty in the classic model; subtracting them from the incumbent's vote-total gives opposition actors their estimate of the incumbent's ability to win a post-election standoff.

Thus the predicted relationship between manipulation and protest likelihood in the regime-weakness model is positive: more manipulation makes protest more likely for a given vote-share. However, the inverse is also to be expected: as incumbent vote-share increases for any fixed level of fraud, the opposition should perceive a stronger incumbent and be less likely to protest. In sum, the effect of fraud should be conditional on the claimed vote-share [@Rozenas2016b]. This produces a first, baseline hypothesis.

@@

It is important to note that the same hypothesis is also generated if we assume that protest initiation is broadly, rather than narrowly, strategic. The mechanism, however, is different. Here, the argument is that widespread grievances result in low natural popularity for the incumbent; at election time, the incumbent must compensate for this low popularity by manipulating the election. Protest initiation is made more likely both by the population's grievances alone [@Beaulieu2014a;@Brancati2016], and by the coordination device provided by the 'stolen' election [@tucker2007]. As natural vote-share increases for a fixed level of manipulation, the likelihood of protest initiation should decline due to reduced grievances; for a fixed level of popularity, increasing manipulation should increase the likelihood of protest by raising the plausibility of a stolen election among dissatisfied citizens.

We should only expect H1 to be falsified in the event that protest is driven by narrowly strategic concerns *and* election manipulation is indicative of regime strength. In this case, opposition actors should be deterred from launching protest movements as increasing manipulation indicates deeper regime resilience (in the form of resources, weak rule of law, loyalty of agents, and so on). The regime-strength model, as I call it, is defended in more detail below.

## Developing the regime-strength model of electoral manipulation

<!--
Putting it all together: this theory assumes that authoritarian incumbents generally prefer order to unrest, and that protesters---who under either a narrow or broadly strategic view are aiming for some kind of concession from the incumbent---want to demonstrate the popularity of their cause through large, sustained protests. Electoral protests are a form of competition between the incumbent and opposition actors, in which both sides are trying to demonstrate their strength in order to win concessions from---or defeat outright---the other [@Beaulieu2014a]. The outcome of this contest reveals the underlying strengths of the participants. Consequently, governments that quickly defeat protest movements can be considered stronger than those that take weeks or months to prevail, which in turn can be considered stronger than those governments which collapse in the face of mass protest.  Incumbents with the resources and capacities needed to produce widespread election manipulation can draw on those same tools to buy off, repress, or otherwise quash post-election protest. This yields Hypothesis 2.
-->

>>>>

Mass protest is one of the two primary threats to authoritarian regimes [@svolik_politics_2012], and as such they often go to great lengths to prevent mass protest. These efforts include costly (and risky) investments in the repressive apparatus [@Escriba_Folch2013;@draguPreventiveRepressionTwo2019], co-optation of opposition parties [@reuter2015], and large-scale monitoring and censorship schemes [@King2013]. And they often relentlessly hound would-be protest entrepreneurs to deter such behavior [@xuRepressCooptAuthoritarian2021a;@mehrlSecretPoliceOrganizations2024]. Mass protest then, is clearly undesirable from the authoritarian incumbent's point of view, and we can plausibly assume that once an electoral protest movement begins, the incumbent will exert considerable effort to quash it. Governments that are stronger, relative to the protesting opposition, will defeat protest movements faster than will weaker governments.^[In both cases, this assume that incumbent survives the protest movement, by far the modal outcome. When the opposition wins, an outcome considered in some of the empirical models below, the assumption is reversed: a quicker opposition victory indicates a weaker incumbent.] As argued below, electoral manipulation can be demonstrative of capacities that may also be used to defeat protest movements.

An important objection could be raised here. Might not a stronger incumbent, feeling secure in their power, be willing to tolerate opposition protest longer than a weak incumbent? It is true that incumbents do sometimes tolerate protest movements [@yuenNeitherRepressionConcession2017]. However, there are two reasons this is unlikely to be the case systematically for the circumstances studied here. First, authoritarian governments are more likely to tolerate protests aimed at low-salience, local issues [@lorentzenRegularizingRiotingPermitting2013], and more likely to clamp down on protests that have higher concession costs [@kleinDynamicsPoliticalProtests2018]. As electoral protests challenge the incumbent's ostensible victory and right to remain in office, they are directly threatening to the regime; the concession costs of stepping down or holding new elections are very high. Second, authoritarian leaders are working in an information-poor environment [@guriev2020], in which mass protests can unexpectedly grow from small sparks to wildfires that quickly sweep away the regime [@Kuran1991]. "Ignoring" protests can lead to further mobilization against authoritarian regimes [@Bishara2015], as can concessions in the absence of durable political reform [@leuschnerStateConcessionsProtest2024]. We should thus expect incumbents to work to defuse electoral protests as quickly as possible.

The central argument of this manuscript is that, in authoritarian regimes, incumbents who develop the tools necessary to manipulate elections extensively will also have the tools necessary to quickly prevail over protest movements. This proposition builds off Simpser's foundational argument that manipulated elections can send costly signals of the incumbent's "capacities, resources, [and] dispositions" that will help it retain power in the future [@simpser2013, p. 80]---that is, of its political strength. However, I depart from Simpser in two ways. First, by incorporating more recent research on the principal-agent dynamics inherent in election manipulation, I argue that incumbents who can produce large amounts of election manipulation necessarily have "capacities, resources and dispositions" helpful for defeating protests. This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver]. Second, mine is not a signalling argument; good manipulators are likely to have deeper wells of strength than weaker manipulators, whether those resources are successfully signaled to opposition actors or no.

@@

This argument---that successful, large-scale election manipulation draws on the same resources a state may use to suppress protest---turns the deterrent logic on its head. Instead of signalling weakness, large-scale fraud indicates a well of resilience: control over resources, impunity for pro-regime lawbreaking, and loyal agents. As a result, governments that build and deploy an electoral manipulation apparatus are simultaneously inoculating themselves against the expected cost of protest.

```{=latex}
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[]
\begin{tabular}{@{}rll@{}}
\toprule
Incumbent's genuine popularity:       & \multicolumn{2}{c}{Low}  \\ \midrule
\multicolumn{1}{r|}{Protest is assumed to be:}                  & \multicolumn{1}{c|}{Instrumental} & \multicolumn{1}{c|}{Multipurpose} \\ \midrule
\multicolumn{1}{|l|}{Manipulation...} &  & \multicolumn{1}{l|}{} \\ \midrule
\multicolumn{1}{|r|}{...obscures strength / indicates weakness} & \multicolumn{1}{l|}{Increasing}   & \multicolumn{1}{l|}{Increasing}   \\ \midrule
\multicolumn{1}{r|}{...reflects strength}                       & \multicolumn{1}{l|}{Decreasing}   & \multicolumn{1}{l|}{Decreasing}   \\ \bottomrule
\end{tabular}
\caption{Expected relationship between protest duration and increasing election manipulation}
\label{tab:duration}
\end{table}
```

How, then, would this view of electoral manipulation map onto the instrumental vs. multipurpose views of protest discussed above? While regime-weakness models have exclusively (to my knowledge) been framed around protest initiation, it is straightforward to extend their predictions to protest duration. Regime-weakness models of electoral manipulation treat election manipulation and genuine votes as substitute goods. For any given vote-share, an increase in manufactured votes reduces the number of genuine votes; that is, the higher the number of manufactured votes for any given vote total, the less likely the incumbent is to survive a challenge. In an instrumental mode of protest, strategic opposition actors perceive the incumbent's relative lack of popularity, and count on the majority of the public to back their challenge. Increasing manipulation indicates deeper unpopularity (holding vote-share constant), so we should expect a less popular incumbent to fail more quickly. The same logic holds, with slightly different mechanisms, if protests are assumed to be multipurpose. Here, greater manipulation is used to overcome deeper unpopularity (i.e. widespread grievances); more manipulation indicates an incumbent facing an increasingly powerful challenger. Thus, the regime-weakness model of manipulation should predict that the duration of unsuccessful protests will be increasing in election manipulation: more manipulation indicates a stronger challenger, which will take longer to defeat on average.

For the regime-strength model of manipulation, the opposite relationship is expected, as laid out above. Comparing Table \@ref(tab:duration) and \@ref(tab:initiation) shows that the hypotheses generated by the preferred model--one in which manipulation indicates strength and protest is multipurpose are distinct from those in other quadrants.

Need to discuss vote-share next:

@@
